The Magnetorotational Instability Prefers Three Dimensions

### 2019-08-19 paper/paper.tex

While surprising, preferred three-dimensional modes can be predicted analytically.
We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming where $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$, 
%\begin{widetext}
\begin{equation}
  \label{eq:asymp}
\sigma^{2} =  -\frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( \frac{d^2}{\pi^{2}} k_z^2 - R^{2}\right)  -  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
\end{equation}
where
\begin{equation}
\Upsilon \ = \ \frac{\left(6+\pi ^2\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
\end{equation}
%\end{widetext}
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is negative definite and leads ultraviolet divergence in the absence of higher-order effects.

>>>>

While surprising, preferred three-dimensional modes can be predicted analytically.
We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) with the shear asymptotically close to the critical two-dimensional rate, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $S = S_c - \epsilon \sqrt{R}$, $k_{z} \sim \mathcal{O}(\epsilon)$, $\sigma \sim k_{y} \sim \mathcal{O}(\epsilon^{2})$, 
%\begin{widetext}
\begin{equation}
  \label{eq:asymp}
\sigma^{2} =  \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left( R^{2} - \frac{d^2}{\pi^{2}} k_z^2 \right)  +  \Upsilon \, k_{y}^{2}  \right] +  \ldots,
\end{equation}
where
\begin{equation}
\Upsilon \ = \ \frac{\left(6+\pi ^2\right) q+\pi^2-6}{12 }\  \approx \ 1.31 \ \  \text{for}  \ \  q = \frac{3}{4}.
\end{equation}
%\end{widetext}
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is negative definite and leads ultraviolet divergence in the absence of higher-order effects.
